### 2019-11-27 2-Chapters/5-Chapter/chapter5.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

\abstractStartChapter{}%
%
% https://english.stackexchange.com/questions/35472/approach-to-or-approach-for
% approach to ? approach for ? approach of ?
In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
in the previous Chapter~\ref{chapter:4}.
We restrict to the easier case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$,
%  because it was found very hard to analyze the previously studied IoT network models,
under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
%
We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature, and we define the centralized regret as a measure of performance of multi-player bandits algorithms.
%
Using a decomposition of the centralized system regret, we start by explaining the intuition about the expected behavior of any efficient decentralized algorithm, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
Combining them with the \klUCB{} index policy gives efficient algorithms, that achieve state-of-the-art performance in terms of their regret,
which is also backed up by numerical simulations.
We conclude by reviewing different extensions of the multi-players models.

@@

The first idea comes from \cite{Anandkumar11}, and the intuition behind it is quite simple, even if the mathematical derivations are not.
All players will build an estimate $\widehat{M}^j(t)$ of the number of player, that start by $\widehat{M}^j(0)=1$. As soon as one collision is observed, a player knows that $M\geq2$, so $\widehat{M}^j(t+1) = 2$.
Then, based on probabilistic computations on the expected number of collisions if there were $m$ players, all following the same strategy (\eg, \RhoRand{} in the case of ), and because the formula is simple to compute for different $m$, the authors proposed a statistical test of the hypothesis $M \leq \widehat{M}^j(t)$ against $M > \widehat{M}^j(t)$ which is expressed as a simple comparison of the current number of collisions (since last update of $\widehat{M}^j(t)$) against a threshold.
If a player observed ``too many'' collisions (that are unlikely to be caused by only $\widehat{M}^j(t) - 1$ other players), then she increases her current estimate $\widehat{M}^j(t+1) = \widehat{M}^j(t) + 1$.

@@

Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of player.
We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for \RandTopM{} and \MCTopM.
Building from the theoretical analysis given for \rhoRandEst{} in \cite{Anandkumar11}, and the analysis of \MCTopM-\klUCB, we believe it is possible to show that the aforementioned extension also achieves sub-linear regret without requiring players to know $M$ in the beginning of the bandit game.
More precisely, we believe that \MCTopM-\klUCB{} can still give order-optimal logarithmic regret if $M$ players use it, and this extension of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is not included due to space constraints.
Writing its proof and performing more simulations are left as possible future work.

@@

Until the very recent work of \cite{KaufmannAbbas19}, it was unknown if a logarithmic regret was possible.
They proposed an algorithm that is based on two phases: first, players will learn $M$ and learn orthogonal ranks, using the ``communication trick'' of~\cite{BoursierPerchet18}, and then one player is elected as a leader and the others are followers.
This second step, after initialization, is also using epoch of increasing lengths $2^p$.
Until the optimal solution, each epoch is alternating three phases.
Followers start by a Round-Robin uniform exploration with no collision, then they use the ``communication trick'' to communicate their samples to the leader, and finally they start an exploitation phase until the end of the game if the leader player tells them so.
The leader can thus collect enough samples from all arms and all players, and is able to solve iteratively the stable marriage problem, and sends back the successive estimate of the solution to the players.
%
In the easier case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO\left( M K \log(K T) + K M^3 \log(M K T) / \Delta + K M^2 (\log(M \log(K T) / \Delta))^2 \right)$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{1+\kappa})$ for any positive constant $\kappa$.

>>>>

\abstractStartChapter{}%
%
% https://english.stackexchange.com/questions/35472/approach-to-or-approach-for
% approach to ? approach for ? approach of ?
In this chapter, we are interested in a more formal approach to the decentralized learning problem presented of many end-devices accessing the same IoT wireless network
than in the previous Chapter~\ref{chapter:4}.
We restrict to the easiest case of at most $M \leq K$ devices using a network with $K$ orthogonal frequency channels.
Each device has data to transmit at every instant, that is we restrict to the case of an activation probability of $p=1$,
%  because it was found very hard to analyze the previously studied IoT network models,
under the hypothesis of random activation patterns of Chapter~\ref{chapter:4}.
%
We discuss three feedback levels that give variants of the multi-players MAB model previously studied in the literature, and we define the centralized regret as a measure of performance of multi-player bandits algorithms.
%
Using a decomposition of the centralized system regret, we start by explaining the intuition about the expected behavior of any efficient decentralized algorithm, and then we propose \RandTopM{} and \MCTopM, two new orthogonalization schemes.
Combining them with the \klUCB{} index policy gives efficient algorithms, that achieve state-of-the-art performance in terms of their regret,
which is also backed up by numerical simulations.
We conclude by reviewing different extensions of the multi-players models.

@@

The first idea comes from \cite{Anandkumar11}, and the intuition behind it is quite simple, even if the mathematical derivations are not.
All players will build an estimate $\widehat{M}^j(t)$ of the number of players, that start by $\widehat{M}^j(0)=1$. As soon as one collision is observed, a player knows that $M\geq2$, so $\widehat{M}^j(t+1) = 2$.
Then, based on probabilistic computations on the expected number of collisions if there were $m$ players, all following the same strategy (\eg, \RhoRand{} in the case of ), and because the formula is simple to compute for different $m$, the authors proposed a statistical test of the hypothesis $M \leq \widehat{M}^j(t)$ against $M > \widehat{M}^j(t)$ which is expressed as a simple comparison of the current number of collisions (since last update of $\widehat{M}^j(t)$) against a threshold.
If a player observed ``too many'' collisions (that are unlikely to be caused by only $\widehat{M}^j(t) - 1$ other players), then she increases her current estimate $\widehat{M}^j(t+1) = \widehat{M}^j(t) + 1$.

@@

Similarly to what was proposed for the \rhoRandEst{} algorithm in \cite{Anandkumar11}, we could have worked on proposing and analyzing an extension of our proposals that could efficiently learn the value of $M$ the number of players.
We implemented this \rhoRandEst{} policy in our library SMPyBandits \cite{SMPyBandits}, as well as this mechanism for \RandTopM{} and \MCTopM.
Building from the theoretical analysis given for \rhoRandEst{} in \cite{Anandkumar11}, and the analysis of \MCTopM-\klUCB, we believe it is possible to show that the aforementioned extension also achieves sub-linear regret without requiring players to know $M$ in the beginning of the bandit game.
More precisely, we believe that \MCTopM-\klUCB{} can still give order-optimal logarithmic regret if $M$ players use it, and this extension of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is not included due to space constraints.
Writing its proof and performing more simulations are left as possible future work.

@@

Until the very recent work of \cite{KaufmannAbbas19}, it was unknown if a logarithmic regret was possible.
They proposed an algorithm that is based on two phases: first, players will learn $M$ and learn orthogonal ranks, using the ``communication trick'' of~\cite{BoursierPerchet18}, and then one player is elected as a leader and the others are followers.
This second step, after initialization, is also using epoch of increasing lengths $2^p$.
Until the optimal solution, each epoch is alternating three phases.
Followers start by a Round-Robin uniform exploration with no collision, then they use the ``communication trick'' to communicate their samples to the leader, and finally they start an exploitation phase until the end of the game if the leader player tells them so.
The leader can thus collect enough samples from all arms and all players, and is able to solve iteratively the stable marriage problem, and sends back the successive estimate of the solution to the players.
%
In the easiest case when their is a unique optimal matching, their algorithm Multiplayer Explore-Then-Commit (M-ETC) is the first one to achieve logarithmic regret, in the form of $R_T = \cO\left( M K \log(K T) + K M^3 \log(M K T) / \Delta + K M^2 (\log(M \log(K T) / \Delta))^2 \right)$, if $\Delta$ is defined as the gap between the utility of the best matching and the utility of that of the matching with second best utility.
In the generic case, their algorithm M-ETC with Elimination achieves a regret of $\cO((\log(T))^{1+\kappa})$ for any positive constant $\kappa$.

### 2019-11-27 2-Chapters/6-Chapter/nonstatbandits/NonStatB.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

A \emph{piece-wise stationary bandit model} generalizes the model of Chapter~\ref{chapter:2}.
It is characterized by a set of $K$ arms, and a horizon $T$.
A (random) stream of rewards $(Y_{k,t})_{t\in[T]}$ is associated to each arm $k \in [K]$.
We assume that the rewards are bounded, and wlog we assume that $Y_{k,t} \in [0,1]$.
We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on her past observations, and receives the corresponding reward $r(t) \eqdef Y_{A(t),t}$.
At time $t$, we denote by $k_t^*$ an optimal arm, \ie, an arm with maximal expected reward $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$ (possibly not unique).

@@

\paragraph{Two example of problems.}\label{par:6:benchmark1}
%
To illustrate the model, we give here two examples of piece-wise stationary problems.
% problems of our benchmark,
used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.

@@

\textbf{Problem 3.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
At every break-points, almost all arms change,
and means are bounded in $[0.01, 0.07]$.
The gaps $\Delta$ are much smaller than for the first problems, with amplitudes ranging from $0.02$ to $0.001$.
Note that the assumptions of the regret upper bounds for \GLRklUCB{} are violated, as well as the assumptions for the analysis of \MUCB{} and \CUSUMUCB.
It is interesting to say that this problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.
%
The details fall outside of the scope of this section, and are well explained in the paper \cite{CaoZhenKvetonXie18}.

>>>>

A \emph{piece-wise stationary bandit model} generalizes the model of Chapter~\ref{chapter:2}.
It is characterized by a set of $K$ arms, and a horizon $T$.
A (random) stream of rewards $(Y_{k,t})_{t\in[T]}$ is associated to each arm $k \in [K]$.
We assume that the rewards are bounded, and without loss of generality we assume that $Y_{k,t} \in [0,1]$.
We denote by $\mu_k(t) \eqdef  \bE[Y_{k,t}]$ the mean reward of arm $k$ at round $t$.
At each round $t$, a decision maker has to select an arm $A(t)\in[K]$, based on her past observations, and receives the corresponding reward $r(t) \eqdef Y_{A(t),t}$.
At time $t$, we denote by $k_t^*$ an optimal arm, \ie, an arm with maximal expected reward $\mu_{k_t^*}(t) \eqdef \max_k \mu_k(t)$ (possibly not unique).

@@

\paragraph{Two examples of problems.}\label{par:6:benchmark1}
%
To illustrate the model, we give here two examples of piece-wise stationary problems
% problems of our benchmark,
used for the numerical experiments presented in Section~\ref{sec:6:NumericalExperiments}.

@@

\textbf{Problem 3.} (see Figure~\ref{fig:6:Problem_5}) This problem is harder, with $K=6$, $\Upsilon_T=8$ and $T=20000$.
At every break-point, almost all arms change,
and means are bounded in $[0.01, 0.07]$.
The gaps $\Delta$ are much smaller than for the first problems, with amplitudes ranging from $0.02$ to $0.001$.
Note that the assumptions of the regret upper bounds for \GLRklUCB{} are violated, as well as the assumptions for the analysis of \MUCB{} and \CUSUMUCB.
It is interesting to say that this problem is inspired from Figure~3 of \cite{CaoZhenKvetonXie18}, where the synthetic data was obtained from manipulations on a real-world database of clicks from \emph{Yahoo!}.
%
The details fall outside of the scope of this section, and are well explained in the paper \cite{CaoZhenKvetonXie18}.

### 2019-11-27 2-Chapters/Conclusion/conclusion.tex
I fixed a lot of small typing mistakes, reported by Nathalie Mitton.
- Thanks to her!
- Shame on me haha

Our main contribution to answer our problematic is a model of Internet of Things networks, based on an ALOHA-like protocol slotted in time and frequency, and that considers one IoT base station serving a large number of devices following the IoT constraints (low-cost, low duty cycle, long range etc).
In such IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacities.
%  (limited hardware)
We proposed two models, whether the considered IoT standard enforces the devices to try to retransmit a few times their up-link packet in case of a failed transmission (\ie, a collision with other devices),
or to drop their packet after a failed transmission (\ie, with or without retransmission).
Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order that the numerous IoT devices improve their spectrum access scheme, on their own, by learning which of the $K$ frequency channels are the less occupied ones by the environmental traffic (in average).

>>>>

Our main contribution to answer our problematic is a model of Internet of Things networks, based on an ALOHA-like protocol slotted in time and frequency, and that considers one IoT base station serving a large number of devices following the IoT constraints (low-cost, low duty cycle, long range etc).
In such a IoT network, we focused on many dynamically reconfigurable IoT end-devices, that are able to run low-cost and low-complexity decision making algorithms, embedded on each device using their limited computational and storage capacities.
%  (limited hardware)
We proposed two models, whether the considered IoT standard enforces the devices to try to retransmit a few times their up-link packet in case of a failed transmission (\ie, a collision with other devices),
or to drop their packet after a failed transmission (\ie, with or without retransmission).
Both models are realistic, and for both cases we proposed to use multi-armed bandit (MAB) algorithms, such as \UCB{} or Thompson sampling, in order that the numerous IoT devices improve their spectrum access scheme, on their own, by learning which of the $K$ frequency channels are the less occupied ones by the environmental traffic (in average).
